---
title: "Project 3"
author: "Lakshmi Mounica Marupeddi"
date: "2024-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Projecr 3

### Import Libraries 
```{r}
#install.packages("nycflights13")
library(nycflights13)
library(dplyr)

```

### Question 1
What is the number of rows in the "flights" set?
```{r}
nrow(flights)
```

### Question 2
What is the number of character variables in the "flights" set? (Just check the dataset as-is, do not make any changes)
```{r}
str(flights)
```
### Question 3
How many unique destination airports are there in flights data?
```{r}

flights %>% 
  distinct(dest) %>%
  summarize(uniqueDestinations = n())
```
### Question 4
 How many unique origin airports are there in the flights data?
```{r}
#
flights %>% 
  distinct(origin) %>% 
  summarize(uniqueOriginAirports = n())


```
### Question 5
Which airports are among the origin airports?
```{r}
# 
flights %>% 
  distinct(origin)
```

### Question 6
From the LGA airport, how many unique destination airports are being flown to?
```{r}
# 
flights %>% 
  filter(origin == "LGA") %>% 
  distinct(dest) %>% 
  summarize(UniqueDestsFromLGA = n())
```
### Question 7
How many flights have originated from "LGA" airport?
```{r}
# 
flights %>% filter(origin == "LGA") %>% count()

```
### Question 8
How many unique carriers have flown from "JFK"?
```{r}
# 
flights %>% 
  filter(origin == "JFK") %>% 
  distinct(carrier) %>% 
  summarize(uniqueCarrriersFromJFK = n())
```
### Question 9
How many trips did the carrier "AA" make in March?
```{r}
# 
flights %>% 
  filter(carrier == "AA" & month == 3) %>% 
  summarize(MarchTripsByAA = n())
```
### Question 10
Find the carrier that has flown the most in May 2013.
```{r}
# 
flights %>% 
  filter(year == 2013 & month == 5) %>%
  group_by(carrier) %>% summarize(MaxTripsCarriers = n()) %>% 
  filter(MaxTripsCarriers == max(MaxTripsCarriers))
```
### Question 11
What is the carrier "UA"s second most popular destination in May 2013?
```{r}
# 
temp = flights %>% filter(year == 2013 & month == 5 & carrier == "UA") %>% group_by(dest) %>% summarize(totalFlights = n()) %>% arrange(-totalFlights)
temp[2,]
```
### Question 12
Now check the "airports" data. This data contains information on the airports. Which variables are part of the airports dataset?
```{r}
# 
glimpse(airports)
```
### Question 13
What is the name of the airport coded as "1H2"? Hint: Check the "faa" column in airports data.
```{r}
# 
airports %>% 
  filter(faa == "1H2") %>% 
  select(name)
```
### Question 14
Which month has the smallest mean total delay in the fl_new data?
```{r}
# Create a new column in the flights set called "total_delay". Select only the month, carrier, origin, dest, distance, arr_delay columns. Save this to a new dataset called fl_new.
fl_new = flights %>% mutate(total_delay = dep_delay+arr_delay) %>% select(month, carrier, origin, dest, distance,total_delay,  arr_delay) 

# smallest mean
fl_new = flights %>% 
            group_by(month) %>% 
            summarize(smallestMean = mean(dep_delay+arr_delay,  na.rm = TRUE)) %>% 
            arrange(smallestMean) 
fl_new[1, ]
```

